Design and Implementation of Probabilistic Programming Language Anglican

### 2017-02-25 paper/paper.tex
small edits to section 8

A probabilistic program in Anglican may look almost like a
Clojure program. However, the purpose of executing a
probabilistic program is different from that of a `regular'
program: instead of producing the result of a single execution,
a probabilistic program computes, exactly or approximately, the
distribution from which execution results are drawn.  Computing
the distribution is facilitated by an inference algorithm.

@@

Consider the following simplified Anglican query:

@@

(defquery baz
  (foo))
\end{lstlisting}

An execution of the query may result in the following
sequence of checkpoint invocations:
\begin{lstlisting}[style=default]
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C2 ...)
  (sample 'C1 ...)
  (sample 'C1 ...)
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C3 ...)
\end{lstlisting}
According to the addressing scheme, the addresses generated
for these invocations are
\begin{lstlisting}[style=default]
[C1 0][C2 0][C2 1][C1 16][C1 17][C1 18][C2 16][C3 0]
\end{lstlisting}
If the program is run by a Metropolis-Hastings algorithm, then a
small change usually takes place in the sequence of checkpoints
with each invocation, and the new sequence may become
\begin{lstlisting}[style=default]
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C1 ...)
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C2 ...)
  (sample 'C3 ...)
\end{lstlisting}
The addresses for the new sequence are
\begin{lstlisting}[style=default]
  [C1 0][C2 0][C1 16][C1 17][C2 16][C2 17][C3 0]
\end{lstlisting}
Despite the change, the correspondence between checkpoints of
similar types occurring in similar positions (relative positions
in contiguous subsequences of a certain type) is preserved, and
drawn values can be reused efficiently:

@@

This way, each occurrence of a checkpoint has unique address,
but small disturbances --- removal or addition of a single or just
a few checkpoints --- are unlikely to derail the entire sequence.
The probability of derailment depends on the padding. The
padding can be safely, and without any impact on performance,
set to rather large numbers. However, rounding up to a multiple
of 16 proved to be appropriate for all practical purposes.
\iftoggle{full}{%

>>>>

A probabilistic program in Anglican may look almost like a
Clojure program. However, the purpose of executing a
probabilistic program is different from that of a `regular'
program: instead of producing the result of a single execution,
a probabilistic program computes, exactly or approximately, the
distribution from which execution results are drawn.  Characterising
this distribution is done by the inference algorithm.

@@

\noindent Consider the following simplified Anglican query:

@@

(defquery baz
  (foo))
\end{lstlisting}
An execution of the query may result in the following
sequence of checkpoint invocations:
\begin{lstlisting}[style=default]
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C2 ...)
  (sample 'C1 ...)
  (sample 'C1 ...)
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C3 ...)
\end{lstlisting}
According to the addressing scheme, the addresses generated
for these invocations are
\begin{lstlisting}[style=default]
[C1 0][C2 0][C2 1][C1 16][C1 17][C1 18][C2 16][C3 0]
\end{lstlisting}
If the program is run by a Metropolis-Hastings algorithm, then a
small change usually takes place in the sequence of checkpoints
with each invocation, and the new sequence may become
\begin{lstlisting}[style=default]
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C1 ...)
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C2 ...)
  (sample 'C3 ...)
\end{lstlisting}
The addresses for the new sequence are
\begin{lstlisting}[style=default]
  [C1 0][C2 0][C1 16][C1 17][C2 16][C2 17][C3 0]
\end{lstlisting}
Despite the change, the correspondence between checkpoints of
similar types occurring in similar positions (relative positions
in contiguous subsequences of a certain type) is preserved, and
drawn values can be reused efficiently:

@@

This way, each occurrence of a checkpoint has a unique address,
but small disturbances --- removal or addition of a single or just
a few checkpoints --- are unlikely to derail the entire sequence.
The probability of derailment depends on the padding. The
padding can be safely, and without any impact on performance,
set to rather large numbers. However, rounding up to a multiple
of 16 proved to be appropriate for all practical purposes.
\iftoggle{full}{%
